Ligeti, Stockhausen and Koenig:\\Aesthetics of Electronic Music

### 2015-09-23 manuscript.tex
Clearing and reordering existing Stockhausen ideas

\begin{MyShadequote}
  On several previous occasions, when I have been asked to explain the composition of electronic music, I have described four characteristics that seem important to me for electronic composition as distinguished from the composition of instrumental music:
\end{MyShadequote}

The titles of the next subsections, \ref{subs:stockhausen:time} -- \ref{subs:stockhausen:noise}, were copied directly from Stockhausen's paper.
In each subsection I add clarification, based on the rest of the paper, as needed.

\subsubsection{The correlation of the coloristic, harmonic-melodic, and metric-rhythmic aspects of composition}
\label{subs:stockhausen:time}

Most of the current paper by Stockhausen deals with this element.
In my opinion, Stockhausen attitude toward the correlation between different compositional aspects could be described by two ideas.

\begin{itemize}
  \item He classifies different sonic elements as different perceptions of the time domain.
  For example, periodic elements of short time intervals are intercepted as pitch, and therefore as the basis for melody and harmony;
  Periodic elements of medium time interval are intercepted as tempo and rhythm;
  Those of larger time intervals supports the musical structure.
  Furthermore, according to Stockhausen these ``time spans'' range across 7 octaves:
  Pitch spread across the 7 octaves of the piano;
  Tempo and rhythm range from 1/16 of a second up to 8 seconds;
  Lastly, time spans of 8 to 1024 seconds are intercepted as structural building blocks.

  \item Based on the above, Stockhausen approaches to electronic music as a tool for total serializm (TODO: {total serializm reference}).
  Furthermore, he argues that serializm must exist in electronic composition, inherently.
  The thinking path that led Stockhausen to this argument is clear, but it is still hard for me to agree with him.
  Later, I will try to confront this idea with those reflected in his and Ligeti's music.
\end{itemize}

\subsubsection{The composition and de-composition of timbres}
\label{subs:stockhausen:timber}

\subsubsection{The characteristic differentiation among degrees of intensity}
\label{subs:stockhausen:intensity}

I'm not sure I fully understand the meaning of this element, not from the current paper, at least.
I will leave this element ``out of scope'' for now.

\subsubsection{The ordered relationships between sound and noise}
\label{subs:stockhausen:noise}

main area to compare Ligeti and Stockhausen ideas TODO.

>>>>

\begin{MyShadequote}
  On several previous occasions, when I have been asked to explain the composition of electronic music, I have described four characteristics that seem important to me for electronic composition as distinguished from the composition of instrumental music:

  \begin{itemize}
    \item The correlation of the coloristic, harmonic-melodic, and metric-rhythmic aspects of composition
    \item The composition and de-composition of timbres
    \item The characteristic differentiation among degrees of intensity
    \item The ordered relationships between sound and noise
  \end{itemize}

  Here, I would like to discuss only the correlation of timbre, pitch, intensity, and duration (p. 39).
\end{MyShadequote}

Stockhausens's paper deals mainly with the 1\textsuperscript{st} point above.
However, I will try to glean additional characteristics of electronic music from the paper as well.

\subsubsection{The correlation of timbre, pitch, intensity, and duration}
\label{subs:stockhausen:time}

Stockhausen classifies different sonic elements as different perceptions of the time domain, thus suggesting a unified time model electronic music.
Periodic elements of short time intervals are intercepted as pitch, and therefore as the basis for melody and harmony;
Periodic elements of medium time interval are intercepted as tempo and rhythm;
Those of larger time intervals supports the musical structure.
Furthermore, according to Stockhausen these ``time spans'' range across 7 octaves each:
Pitch spread across the 7 octaves of the piano;
Tempo and rhythm range from 1/16 of a second up to 8 seconds;
Lastly, time spans of 8 to 1024 seconds are intercepted as structural building blocks.

Based on the above, Stockhausen approaches electronic music as a tool for total serializm (TODO total serializm reference).
Furthermore, he argues that serializm must exist in electronic composition, inherently.
